%Input preamble
\input{preamble}
\let\counterwithout\relax
\let\counterwithin\relax
\definecolor{maroon}{HTML}{4B0082}

\begin{document}
%\onehalfspacing

\noindent \textbf{Properties of OLS.}\\
\noindent Jorge Luis García \\
\noindent e-mail: jlgarci@clemson.edu\\

\noindent Recall that the (sample) model is
\begin{equation}
\bm{y} = \bm{x} \cdot \bm{\beta} + \bm{e}, \label{eq:model}
\end{equation}
\noindent and that the estimated (OLS) residual vector is 
\begin{align}
\hat{\bm{e}} = \bm{y} - \bm{x} \cdot \hat{\bm{\beta}}.	
\end{align}

\noindent \textbf{Projection and Annihilation Matrices.} The residual is ``all of the variation'' in $\bm{y}$ not explained in $\bm{x}$: 
\begin{align}
	\hat{\bm{e}} & = \bm{y} - \bm{x} \cdot {\left( \bm{x}'\bm{x} \right)}^{-1} \bm{x}'\bm{y} \nonumber \\ 
	 			 & = \left[ I_N - \bm{x} {\left( \bm{x}'\bm{x} \right)}^{-1} \bm{x}' \right] \bm{y} \nonumber \\ 
	 			 & := \left[ I_{N} - P_x \right] \bm{y} \nonumber \\
	 			 & := M_x \bm{y}, 
\end{align}
\noindent where $P_x$ is the projection matrix and $M_{x}$ is the annihilation matrix or ``residual maker.'' Multiplying $P_{x}$ by $\bm{y}$ yields $ \bm{x} \cdot \bm{\beta}$ under \textbf{Exogeneity} (the projection of $\bm{y}$ on $\bm{x}$; the part of $\bm{y}$ that $\bm{x}$ explains). Multiplying $M_{x}$ by $\bm{y}$ yields $\hat{\bm{e}}$ (the multiplication makes the residuals; it yields the part of $\bm{y}$ that $\bm{x}$ does not explain).\\

\noindent \textbf{Properties of OLS.} The main properties of OLS are:
\begin{enumerate}
	\item The first order conditions of the OLS problem are
		\begin{align}
			\bm{x}' \bm{x} \hat{\bm{\beta}} - \bm{x}' \bm{y} & = \bm{0} \nonumber \\\
			\bm{x}' \hat{\bm{e}} & = \bm{0}.
		\end{align}
\noindent OLS ``enforces'' the empirical counterpart of the \textbf{Exogeneity} assumption. 
	\item From 1. and the standard specification of the first column of $\bm{x}$ as $\bm{x}_1 = \left[ 1, \cdots, 1 \right]$, $\sum \limits _{i \in \mathcal{I}} e_{i} = 0$. The sum of residuals, and therefore the average of the residuals, is $0$. 
	\item From 1. and 2., $\frac{1}{N} \cdot \sum \limits _{i \in \mathcal{I}} e_{i} = 0$. This implies that $\bar{y} = \bar{\bm{x}} \cdot \bm{\hat{\beta}}$. That is, the (OLS) prediction for the average observation is equal to the average of the dependent variable.
	\item From 1. and 2., it also follows that the average prediction is equal to the average of the dependent variable. 
\end{enumerate}

\noindent \textbf{Partitioned Regression.} Let $k_1$ and $k_2$ be less than $K$ and such that $k_1 + k_2 = K$. The sample model can be written as
\begin{align}
	\bm{y} = \bm{x}_1 \cdot \bm{\beta}_1 + \bm{x}_2 \cdot \bm{\beta}_2 + \bm{e}, 
\end{align}
\noindent where $\bm{x}:= \begin{bmatrix} \bm{x}_1 & \vdots & \bm{x}_2 \end{bmatrix}$ and $\bm{\beta} = \begin{bmatrix} \bm{\beta}_1 \\ \ldots \\ \bm{\beta}_2 \end{bmatrix}$. 

\noindent The first order conditions of the OLS problem are 
\begin{align}
	\begin{bmatrix}
		\bm{x}_1' \bm{x}_1 & \bm{x}_1' \bm{x}_2 \\
		\bm{x}_2' \bm{x}_1 & \bm{x}_2' \bm{x}_2 
	\end{bmatrix} \cdot
	\begin{bmatrix}
		\bm{\hat{\beta}}_1 \\
		\bm{\hat{\beta}}_2	
	\end{bmatrix} = 
	\begin{bmatrix}
		\bm{x}_1 \bm{y} \\ 
		\bm{x}_2 \bm{y}
	\end{bmatrix}, 
\end{align}

\noindent which yield a system to two (vectorial) equations with two unknown vectors: 
\begin{align}
	\bm{x}_1' \bm{x}_1 \bm{\hat{\beta}}_1 + \bm{x}_1' \bm{x}_2 \bm{\hat{\beta}}_2 & =  \bm{x}_1' \bm{y} \nonumber \\
		\bm{x}_2' \bm{x}_1 \bm{\hat{\beta}}_1 + \bm{x}_2' \bm{x}_2 \bm{\hat{\beta}}_2 & =  \bm{x}_2' \bm{y}. 
\end{align}

\noindent Solving, 
\begin{align}
		\hat{\bm{\beta}}_1 & = {\left( \bm{x}_1' M_{\bm{x}_2} \bm{x}_1 \right)}^{-1} \bm{x}_1' M_{\bm{x}_2} \bm{y} \nonumber \\
		\hat{\bm{\beta}}_2 & = {\left( \bm{x}_2' M_{\bm{x}_1} \bm{x}_2 \right)}^{-1} \bm{x}_2' M_{\bm{x}_1} \bm{y}
\end{align}
\noindent where $M_{\bm{x}_\ell}:= \left[ I_{N} - \bm{x}_\ell {\left( \bm{x}_\ell' \bm{x}_\ell \right)}^{-1} \bm{x}_\ell' \right]$ for $\ell = 1,2$ (note that $M_{\bm{x}_\ell}$ is idempotent).\\ 

\noindent This illustrates the most fundamental mechanics of OLS. To estimate $\hat{\bm{\beta}}_1$, OLS ``cleans'' $\bm{x}_1$ and $\bm{y}$ from $\bm{x}_2$ and ``estimates'' OLS with the ``cleaned versions.'' To clarify this further, let $\bm{x}_1^*:= M_{\bm{x}_2} \bm{x}_1$, $\bm{x}_2^*:= M_{\bm{x}_1} \bm{x}_2$, $\bm{y}_1^*:= M_{\bm{x}_2} \bm{y}$, and $\bm{y}_2^*:= M_{\bm{x}_1} \bm{y}$ and note that
\begin{align}
		\hat{\bm{\beta}}_1 & = {\left( {\bm{x}_1^*}' \bm{x}_1^* \right)}^{-1}{\bm{x}_1^*}'\bm{y}_1^* \nonumber \\
		\hat{\bm{\beta}}_2 & = {\left( {\bm{x}_2^*}' \bm{x}_2^* \right)}^{-1}{\bm{x}_2^*}'\bm{y}_2^*. 
\end{align}

\end{document}
