%Input preamble
\input{preamble}
\let\counterwithout\relax
\let\counterwithin\relax
\definecolor{maroon}{HTML}{4B0082}

\begin{document}
%\onehalfspacing

\noindent \textbf{Revisiting Inference in the Classic (Normal) Linear Regression Model.}\\
\noindent Jorge Luis García \\
\noindent e-mail: jlgarci@clemson.edu\\

\noindent \textbf{Likelihood Ratio Test.} Let $\bm{\theta}$ be a parameter vector of interest. A hypothesis is
\begin{align}
	H_0 & : r \left( \bm{\theta} \right) = 0 \nonumber \\
	H_1 & : r \left( \bm{\theta} \right) \neq 0 \nonumber,
\end{align}
\noindent where $r$ is a function that maps a $k$-dimensional argument into a $q$-dimensional space. This function allows distinguishing between an unrestricted estimate and a restricted estimate (which imposes $H_0$). The likelihood ratio is
\begin{align}
	\lambda & = \frac{ L \left( \bm{\hat{\theta}}_r \right) } { L \left( \bm{\hat{\theta}}_u \right) } \nonumber \\ 
		    & \leq 1, 
\end{align}
\noindent where $L \left( \cdot \right)$ is the sample likelihood. It is out of the scope of this class to show that 
\begin{align}
	- 2 \ln \lambda \sim {\chi^2}_{(q)}, 
\end{align}
\noindent which enables testing the hypothesis. This likelihood-ratio test uses both estimates, the restricted and the unrestricted.\\

\noindent \noindent \textbf{Wald Test.} An alternative for testing $H_0$ is using a Wald test. The statistic (previously discussed) is 
\begin{align}
	W & = \left[ r ( \bm{\hat{\theta}}_u \right]' \left[ \var \left( r \left( \bm{\hat{\theta}}_u \right) \right) \right] \left[ r ( \bm{\hat{\theta}}_u \right] \nonumber \\ 
	  & \sim {\chi^2}_{(q)}, 
\end{align}

\noindent where $\var \left(  \bm{\hat{\theta}}_u \right)$ is known but $\var \left( r ( \bm{\hat{\theta}}_u \right)$ is not. 

\begin{theorem} \label{delta} (Delta Method) Let $\bm{u}$ be a $k \times 1$ vector with $\bm{u} \sim \mathcal{N} \left( \bm{\mu}, \bm{\Sigma} \right)$. Let $f \left( \bm{u} \right)$ be a function mapping a $k$-dimensional argument to a $q$-dimensional space. Then, $f \left( \bm{u} \right) \sim \mathcal{N} \left( f \left( \bm{\mu} \right) , J \left( \bm{\mu} \right) \bm{\Sigma} J \left( \bm{\mu} \right)' \right)$, where $J \left( \bm{\mu} \right)$ is the Jacobian matrix of $f \left( \cdot \right)$. 
\end{theorem}

\noindent Using Theorem~\ref{delta}, $\var \left( r ( \bm{\hat{\theta}}_u \right) = J \left( \var \left(  \bm{\hat{\theta}}_u \right) \right) \var \left(  \bm{\hat{\theta}}_u \right) J \left( \var \left(  \bm{\hat{\theta}}_u \right) \right)'$. This Wald test only uses unrestricted estimates.

\end{document}