%Input preamble
\input{preamble}
\let\counterwithout\relax
\let\counterwithin\relax
\definecolor{maroon}{HTML}{4B0082}


\begin{document}
%\onehalfspacing

\noindent \textbf{Consistency in the Classic Linear Regression Model.}\\
\noindent Jorge Luis García \\
\noindent e-mail: jlgarci@clemson.edu\\

\noindent \textbf{Probability Limits.} A sequence $\{ x_n \}$ of random variables converges in probability towards the random variable $x$ if, for all $\varepsilon > 0$: 
\begin{align}
	   \plim_{N \rightarrow \infty} | x_{n} - x | & := \lim_{N \rightarrow \infty} \Pr \left( | x_{n} - x | > \varepsilon \right) \nonumber \\ 
									   & = 0. 
\end{align}

\noindent \textbf{The Probability Limit of OLS: The Two-Variable Case.} Recall that the sample model for $i \in \mathcal{I}$ is 
\begin{align}
	y_i = a + b \cdot x_i + e_i. 
\end{align}
\noindent The estimator of b is
\begin{align}
\hat{b} & =  \frac{ \frac{1}{N} \cdot \sum \limits _{i \in \mathcal{I} } \left( x_{i} - \bar{x} \right)  \left( y_{i} - \bar{y} \right) } { \frac{1}{N} \cdot \sum \limits _{i \in \mathcal{I} } { \left( x_{i} - \bar{x} \right) }^2 }, 
\end{align}
which is the sample covariance of $x_i$ and $y_i$ divided by the covariance of $x_i$. It follows that 
\begin{align}
	\plim_{N \rightarrow \infty} \hat{b} = \frac{\cov \left( x_i, y_i \right)} {\var \left( x_i \right)}, 
\end{align}
\noindent a usual and very useful expression to explore the consistency of OLS. For the two-variable case and under \textbf{Exogeneity}: 
\begin{align}
		\plim_{N \rightarrow \infty} \hat{b} & = \frac{\cov \left( x_i, y_i \right)} {\var \left( x_i \right)} \nonumber \\
											 & = \frac{\cov \left( x_i, a + b \cdot x_i + e_i \right)} {\var \left( x_i \right)} \\ 
											 & = \frac{\cov \left( x_i, a \right) + b \cdot \var \left( x_i \right) + \cov \left( x_i,  e_i \right) } {\var \left( x_i \right)} \nonumber \\ 
											 & = b,								 
\end{align}
\noindent which shows that $\hat{b}$ is consistent.\\

\noindent \textbf{The Multivariate Case.} Consider a multivariate sample model for $i \in \mathcal{I}$. The explanatory variables are partitioned into the intercept and two blocks. The first block has one variable, ``the variable of interest.'' The second block has the rest of the variables. The model is
\begin{align}
	y_i = \beta_0 + x_{i1} \cdot \beta_1  + \bm{x}_{i2} \cdot \bm{\beta}_2 + e_i. 
\end{align}

\noindent To clarify, the third term in the model is $x_{i2} \cdot \beta_2 + \cdots + x_{iK} \cdot \beta_K$ (using the notation developed before). The results from partitioned regression indicate that 
\begin{align}
	\plim_{N \rightarrow \infty} \hat{\beta_1} = \frac{\cov \left( x_{i1}^*, y_i \right)} {\var \left( {x_i}^* \right)}. 
\end{align}

\noindent This result is useful: It enables partitioning the model into a block containing a variable of interest, for which consistency could be analyzed, in a simple one-dimensional expression with known properties (using rules for computing variances and covariances). Completing the calculation: 
\begin{align}
	\plim_{N \rightarrow \infty} \hat{\beta_1} & = \frac{\cov \left( x_{i1}^*, y_i \right)} {\var \left( {x_i}^* \right)} \nonumber \\ 
											   & = \frac{\cov \left( x_{i1}^*, \beta_0 + x_{i1} \cdot \beta_1  + \bm{x}_{i2} \cdot \bm{\beta}_2 + e_i \right)} {\var \left( {x_i}^* \right)} \nonumber \\ 
											   & = \beta_1, 
\end{align}
using \textbf{Exogeneity} and $x_{i1}^* \cdot \bm{x}_{i2} \cdot \bm{\beta}_2 = 0$ because $x_{i1}$ is ``clean'' of $\bm{x}_{i2}$ by the fundamental mechanics of OLS. 




\end{document}